---
title: DANL Project
subtitle: "Analyzing Trends in Athlete Salaries Across Countries and Sports"
author: 
  - Name Collin Gates
  - Name Abby Dunnigan
  - Name Meghan Catania
  - Name Teddy Packert
  - Name Emily Peters

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(readxl)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```



# Introduction

About this project :clap:

This project explores the evolution of athlete salaries across different sports and countries over time. We analyze trends, make comparisons, and identify factors influencing these changes

# Data

We utilize a comprehensive dataset of athlete salaries from various sports and countries, spanning multiple years. The dataset includes information such as the year, country, sport, and the corresponding salaries:


## Summary Statistics

```{r}
athlete_salaries <- read_excel("C:/Users/collin/Documents/Data Analytics/project/Most_paid_athletes.xlsx")

```

```{r}
#| results: asis
#| echo: false
rmarkdown::paged_table(athlete_salaries) 
```


```{r}
skim(mpg) %>% 
  select(-n_missing)
```


## MPG and a Type of Cars
The following boxplot shows how the distribution of highway MPG (`hwy`) varies by a type of cars (`class`) :blue_car: :truck: :minibus:.


```{r}
ggplot(data = mpg) +
  geom_boxplot(aes(x = class, y = hwy, fill = class),
               show.legend = F) +
  labs(x = "Class", y = "Highway\nMPG") 
```


